---
title: "Hackathon: Wall Street Journal Sentiment Analysis"
editor: visual
---

## Intro

In a sentiment analysis for the Wall Street Journal for the 2018 to early 2020 time period, we try four different methods for processing sentiment as positive, neutral, or negative over time as based on scores returned from these methods.

## Synergy Report

Our team demonstrated great
synergy for this assignment. The team discussed the plan of action during
class, then the members coordinated virtually. Wen Si and Kiran Jaura generated
the initial concept for the project. They subsequently shared their thoughts
with Glen Cooper and Marcus Sianan, who both consented because they found the
topic/concept to be interesting, important, and relevant to the course. Wen was
most responsible for the design of the product. Kiran put the most effort into
refining and improving upon the structure that Wen devised, and consulted with
Glen and Marcus throughout the process (e.g., ensuring that the visualizations
were as accessible and interpretable as possible for the audience). As team
coordinator, Kiran handled much of the group communication, and was assisted by
Glen and Marcus. Specifically, all members interacted via a series of emails in
order to make sure that everyone was on the same page and that there was
unanimous consent to all modifications of the code. Lastly, Marcus drafted the
synergy report and it was reviewed by all members of the team. Their input was
incorporated into the final draft. Overall, the group had a positive,
collaborative experience. 

## Data Preparation

```{r}


rm(list = ls()) # clear the environment
setwd("C:/Users/19728/Desktop/EPPS 6356/Wen_Files")

# load packages
library(tm)
library(NLP)
#library(austin)
library(ggplot2)
library(readr)
library(tidytext)
library(dplyr)
library(tidyr)
library(lubridate)
library(quanteda)
library(quanteda.textmodels)
library(quanteda.textplots)
library(stm)
library(syuzhet)

#load data

wsj_data <- read_csv("ISDSA_WSJ DATA 1.csv")
doc.corpus <- corpus(wsj_data)
#summary(doc.corpus)

# remove stop words
my.dfm <- dfm(doc.corpus, remove_numbers = TRUE, remove_punct = TRUE, remove_symbols = TRUE, remove = stopwords("english"))
#my.dfm


###### Sentiment Analysis by Time#####
library(readxl)
wsj_data1 <- read_excel("ISDSA_WSJ DATA.xlsx")
wsj_txt <- iconv(wsj_data1$Text)

wsj_sent <- get_nrc_sentiment(wsj_txt)



wsj_szvector = get_sentiment(wsj_data1$Text, method =  "syuzhet")

# get sentiments

wsj_bvector = get_sentiment(wsj_data1$Text, method =  "bing")
wsj_afvector = get_sentiment(wsj_data1$Text, method =  "afinn")
wsj_nrcvector = get_sentiment(wsj_data1$Text, method =  "nrc", lang ="english")


attach(wsj_data1)
library(lubridate)
wsj_data$date = mdy(wsj_data1$entryDate)


new_b<-cbind(wsj_bvector, "Bing")
new_b<-cbind.data.frame(new_b,wsj_data$date)
colnames(new_b)<-c("Score","Method","Date")
new_afv<-cbind(wsj_afvector,"Affin")
new_afv<-cbind.data.frame(new_afv,wsj_data$date)
colnames(new_afv)<-c("Score","Method","Date")
new_nrc<-cbind(wsj_nrcvector,"NRC")
new_nrc<-cbind.data.frame(new_nrc,wsj_data$date)
colnames(new_nrc)<-c("Score","Method","Date")
new_sz<-cbind(wsj_szvector,"Syuzhet")
new_sz<-cbind.data.frame(new_sz,wsj_data$date)
colnames(new_sz)<-c("Score","Method","Date")
new<-rbind(new_b,new_afv,new_nrc,new_sz)
new<-as.data.frame(new)
new$Score<-as.integer(new$Score)

h_line <- mean(new$Score)




```

## Visualization

```{r}
library(ggplot2)

ggplot(new, aes(Date,Score)) + geom_line(aes(color=Method)) + facet_wrap(~Method, ncol = 2) + theme(legend.position="none",axis.text.x = element_text(angle = 90)) + xlab("Date")+ylab("Score") +ggtitle("WSJ Sentiment Scores by Different Sentiment Methods") + scale_color_manual(values = c("Affin" = "goldenrod","Bing"="forestgreen","NRC"="steelblue","Syuzhet"="maroon")) + geom_hline(aes(yintercept=0))
```

In a sentiment analysis for the Wall Street Journal for the 2018 to early 2020 time period, we try four different methods for processing sentiment as positive, neutral, or negative over time as based on scores returned from these methods.

We created the visual to aid in the following objectives:

1\) show baseline of 0 for neutral sentiment to show how methods differ from each other in regards to how distant they are from 0

2\) assist user in identifying which methods show more positive results, which ones show more negative results, and which ones have greater variance vs being packed more tightly together toward 0 (neutral)

We hope this visual will assist sentiment analysis researchers in choosing the right method for their project while understanding how the results can be affected by the method.

Details on sentiment analysis methods in R found here: <https://arxiv.org/pdf/1901.08319>

Note: mean of all Scores for all time periods was slightly positive: 2.25545.
